---
title: "Athenian Demo"
author: "Vojtech Huser"
date: "12/04/2019"
output: 
  html_document: 
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Various uses will be shown by chapter


```{r load}
source('codeToRun.R')
library(DT);library(knitr)

```

# Simple calls of overview functions

Of classes

```{r simple}
get_version()

kable(overview_class())

#dummy

```

# overview_vocab

Overview of vocabularies

```{r ov}

kable(overview_vocab())

```

# overview_vocab and with concept status

Overview of vocabularies with more data on concept status about each vocabulary (look for S and C mostly)

```{r ov2}

kable(overview_vocab_standard())

```




# overview of domains

Overview of 

```{r ov3}

kable(overview_domain())

```




# overview of domains and vocabs

Overview of 

```{r ov4}

kable(overview_domain_and_vocab())

```




# included vocabularies

```{r}
kable(vocabulary)

```


# other

```{r}
#deprecated with no successor in LOINC and RxNorm
search_by_invalid_reason(invalid_reason_value = 'D',vocabulary = c('LOINC','RxNorm')) %>% head()


cid=1710316 #emtricitabine
search_concept_id(cid)

search_relationships(cid,vocabulary = 'RxNorm',full_output = TRUE)  %>% datatable(rownames = FALSE)

#dummy
```
# Maps to relationships use

```{r maps}
#
#translate-dx.csv
tset<-c('ICD10CM','ICD9CM','ICD10')
#start<-concept %>% filter(vocabulary_id %in% tset)
#only taking certain relationship
mt<-concept_relationship %>% filter(relationship_id=='Maps to')

#more info on the concept 1 (from)
mt %<>% left_join(concept, by=c('concept_id_1'='concept_id'))
names(mt)
#restrict to our starting terminologies
mt2<-mt %>% filter(vocabulary_id %in% tset)

mt2 %<>% left_join(concept, by=c('concept_id_2'='concept_id'))
names(mt2)

#all relationships  are valid
mt2 %>% count(invalid_reason.x)

#export
oname='local/translate_dx'
mt2 %>% write_csv(paste0(oname,'-all.csv'))
mt3<-mt2 %>% select(vocabulary_id=vocabulary_id.x,concept_code=concept_code.x,target_concept_id=concept_id_2) 
mt3 %>% write_csv(paste0(oname,'.csv'))
#end
```

# Mapping drugs to ingredients

Clinical Drugs have ingredients. We want a simple file that will list those for all possible clinical drugs.

```{r ingred}
#list all clinical drugs
#for international analysis, also do RxNorm extension
cd<-concept %>% filter(vocabulary_id=='RxNorm') %>% filter(concept_class_id=='Clinical Drug')
#names(concept)
#cd %>% count(concept_class_id)
nrow(cd)
#only taking certain relationship

#looking at what relationship_id to use
 # cid=1710316 #emtricitabine
 # search_concept_id(cid)
 # 
 # search_relationships(cid,vocabulary = 'RxNorm',full_output = TRUE)  %>% datatable(rownames = FALSE)

#filter only those relationships
mt<-concept_relationship %>% filter(relationship_id=='Consists of')

#more info on the concept 1 (from)
mt %<>% left_join(concept, by=c('concept_id_1'='concept_id'))
names(mt)

#restrict to clinical drug list created earlier (note the INNER join now)
mt2<-mt %>% inner_join(cd,by=c('concept_id_1'='concept_id'))

#get explanation on the ingredient
mt2 %<>% left_join(concept, by=c('concept_id_2'='concept_id'))
names(mt2)

#all relationships  are valid
mt2 %>% count(invalid_reason.x)

#export
oname='local/translate_rx_ingredient_component'
mt2 %>% write_csv(paste0(oname,'-all.csv'))
names(mt2)
mt3<-mt2 %>% select(
  #vocabulary_id=vocabulary_id.x,concept_code=concept_code.x,
  concept_code=concept_id_1
  ,target_concept_id=concept_id_2) 
mt3 %>% write_csv(paste0(oname,'.csv'))

#this only does it to Clinical Drug Comp (emtricitabine 200 mg!!) (not ingredient)


#different approach

ing<-concept %>% filter(vocabulary_id=='RxNorm') %>% filter(concept_class_id=='Ingredient')
#names(concept)
#ing %>% count(concept_class_id)
nrow(ing)

names(concept_ancestor)
#restrict
mt<-concept_ancestor %>% inner_join(cd,by=c('descendant_concept_id'='concept_id'))
#restrict to ingredients as ancestors
mt2<-mt %>% inner_join(ing,by=c('ancestor_concept_id'='concept_id'))

#get explanation on the ingredient
#mt2 %<>% left_join(concept, by=c('descendant_concept_id'='concept_id'))
names(mt2)


#export
oname='local/translate_rx_ingredient'
mt2 %>% write_csv(paste0(oname,'-all.csv'))
names(mt2)
mt3<-mt2 %>% select(
  #vocabulary_id=vocabulary_id.x,concept_code=concept_code.x,
  concept_code=ancestor_concept_id
  ,target_concept_id=descendant_concept_id) 
mt3 %>% write_csv(paste0(oname,'.csv'))


```


# Is icd10 subsumed in icd10cm?

Small analysis

```{r inotcm}




i10<-concept %>% filter(vocabulary_id=='ICD10' & invalid_reason=='')
i10cm<-concept %>% filter(vocabulary_id=='ICD10CM'  & invalid_reason=='')
#inner_join(i10$concept_code,i10cm$concept_code) 
#length(intersect(i10$concept_code,i10cm$concept_code) )


nrow(i10)
inInotCM<-setdiff(i10$concept_code,i10cm$concept_code)
nrow(i10cm)
#that many concepts are in Int and not in US version ( ICD10CM)
length(inInotCM)
d<-i10 %>% filter(concept_code %in% inInotCM)
d %>% datatable(rownames = FALSE)
d %>% write_csv('local/inInotCM.csv')
```

 
```{r unmapped, eval=FALSE, include=FALSE}

#Unmapped ICD10CM codes

i10cm<-concept %>% filter(vocabulary_id=='ICD10CM')
i10cm
cid=1326507
search_relationships(cid,full_output = TRUE) %>% View()
mt<-concept_relationship %>% filter(relationship_id=='Maps to')

mt %<>% left_join(concept, by=c('concept_id_1'='concept_id'))
mt %<>% left_join(concept, by=c('concept_id_2'='concept_id'))
names(mt)
mt2<-mt %>% filter(vocabulary_id.x=='ICD10CM')
imap <- mt %>% inner_join(select(i10cm,concept_id), by=c('concept_id_1'='concept_id'))
nomap <- mt %>% anti_join(select(i10cm,concept_id), by=c('concept_id_1'='concept_id'))

concept_relationship %>% filter(relationship_id=) inner_join(i10cm, by)

```




```{r old, eval=FALSE, include=FALSE}

# Old vignette
get_version()

library(tidyverse)
#print all rows for next two commands
options(tibble.print_max = Inf)

overview_class()
overview_vocab()

search_code('250.00')



search_text('itag',vocabulary='RxNorm') 
search_by_invalid_reason(invalid_reason_value = 'D') %>% head()

#deprecated with no successor in LOINC and RxNorm
search_by_invalid_reason(invalid_reason_value = 'D',vocabulary = c('LOINC','RxNorm')) %>% head()

#using full output
search_by_invalid_reason(invalid_reason_value = 'D',vocabulary = c('LOINC','RxNorm'),full_output = TRUE) %>% head()

search_by_invalid_reason(invalid_reason_value = 'D',vocabulary = c('NDC'),full_output = TRUE) %>% head()

#deprecated with sucessor (upgraded)
#search_by_invalid_reason(invalid_reason_value = 'U',vocabulary = 'LOINC',full_output = TRUE) %>% head()



#looking into deprecation by vocabulary_id by year
# first prepare a new column with end year
concept$valid_end_date %<>%  as.character
concept$end_year=str_sub(concept$valid_end_date,1,4)
aa<-search_by_invalid_reason(invalid_reason_value = 'U',full_output = TRUE) %>% count(vocabulary_id,end_year) 
kable(aa)
aa %>% write_csv('local/upgraded.csv')

ab<-search_by_invalid_reason(invalid_reason_value = 'D',full_output = TRUE) %>% count(vocabulary_id,end_year) 
kable(ab)
ab %>% write_csv('local/deprecated.csv')

#reused codes
reu<-search_by_invalid_reason(invalid_reason_value = 'R') 
datatable(reu)
#concept %>% group_by(invalid_reason) %>% count()


  



#extract which LOINC codes appear in mappings

#we need to find all MapsTorelationship that have target terms LOINC codes

#all realtionships
names(concept_relationship)
options(min.print = 100)
options(tibble.print_max = 500)
concept_relationship %>% count(relationship_id) 
aa<-concept_relationship %>% 
  filter(concept_id_1!=concept_id_2) %>% 
  filter(relationship_id=='Maps to') %>%
  left_join(concept,by=c('concept_id_2'='concept_id')) %>% 
  left_join(concept,by=c('concept_id_1'='concept_id')) 
  
names(aa)
ab<-aa %>% filter(vocabulary_id.y=='LOINC') 
#lipid panel, let's see the relationships
cid=2212095
search_relationships(cid)
#we see CPT4 - LOINC eq

ac<-concept_relationship %>% 
  #ignore replationships to self
  filter(concept_id_1!=concept_id_2) %>% 
  filter(relationship_id=='CPT4 - LOINC eq') %>%
  #get explenations of codes
  left_join(concept,by=c('concept_id_1'='concept_id')) %>% 
  left_join(concept,by=c('concept_id_2'='concept_id')) %>% 
  filter(standard_concept.y=='S' )
ac %>% write_csv('CPT4-LOINCeq.csv')


    


# map NDCs to clinical drug (not branded drug)



#filter all relationships of NDCs  
cid=45077293
search_relationships(cid)
names(concept_relationship)
rxnorm <- concept %>% filter(vocabulary_id=='RxNorm')
# mt<-concept_relationship %>% filter(relationship_id=='Maps to')
# mt %<>% left_join(concept, by=c('concept_id_1'='concept_id'))
# mt %<>% left_join(concept, by=c('concept_id_2'='concept_id'))

#map<-mt %>% inner_join(rxnorm %>% select(concept_id), by=c('concept_id_2'='concept_id'))
#map %<>% left_join(concept, by=c('concept_id_1'='concept_id'))
map <- mt %>% filter(vocabulary_id.x=='NDC' & concept_class_id.x=='11-digit NDC')
map %>% write_rds('o:/athena/ndc-map-a.rds')

#if branded drug, I must go further
cid=40171779
cid=1710316

search_relationships(cid,vocabulary = 'RxNorm',full_output = TRUE) %>% head()

#brand drug    #Tradename of   #clinical drug
bcmap <- concept_relationship %>% filter(relationship_id=='Tradename of') %>% rename(cdcid=concept_id_2)

names(mapp)
mapp<-map %>% left_join(bcmap,by=c('concept_id_2'='concept_id_1')) %>% left_join(concept %>% select(concept_id,concept_name),by=c('cdcid'='concept_id'))
names(mapp)
mappp<-mapp %>% mutate(target_cid=if_else(is.na(cdcid),concept_id_2,cdcid))
names(mappp)

mappp %>% write_csv('o:/athena/ndc-map-b-ignore.csv')
#mtcars %>%   mutate(new_col = if_else(mpg*cyl == 126.0, 2, 0)) %>%    head()
mappp %>% select(concept_code=concept_code.x,target_cid) %>% write_rds('o:/athena/ndc-map-b.rds')



#hieararchy for a concept


cid=45077293

cid<-search_code('02HV33Z')$concept_id #%>% unlist()
cid

#atc example
cid<-search_code('J05AF09')$concept_id #%>% unlist()



aa<-search_relationships(cid,full_output = TRUE) 
names(aa)
#View(aa)
#View(aa %>% filter(standard_concept=='C'))

#for icd10 pcs hieararchy the concepts are not marked C

```



# Concept ancestor

```{r ca}
cid=37395835 #familial prostate cancer
#brief output
get_ancestors(cid)  %>% datatable(rownames = FALSE)

#full output (use full output parameter)
```


# Data on single vocabulary_id

```{r vid}
#vid=c('PPI')
vid='RxNorm'
vid='ICD10'
vid='PPI'
#selc<-concept %>% filter(vocabulary_id %in% vid)

selc<-concept %>% filter(vocabulary_id==vid)

#number of concepts
nrow(selc)

selc %>% count(concept_class_id) %>% arrange(desc(n))

selc %>% count(domain_id) %>% arrange(desc(n))

selcr<-concept_relationship %>% inner_join(selc, by=c('concept_id_1'='concept_id'))

#number of relationships
nrow(selcr)
selcr %>% count(relationship_id)%>% arrange(desc(n))

#what it maps to
selcr %<>% left_join(concept,by=c('concept_id_2'='concept_id')) 
#all rels
selcr %>% count(relationship_id,vocabulary_id.y) %>% arrange(vocabulary_id.y,desc(n))

#only  maps to rels
selcr %>% filter(relationship_id=='Maps to') %>% count(vocabulary_id.y,domain_id.y) %>% arrange(desc(n))
```

